NLTK: The Natural Language Toolkit

### 2002-05-14 technical/acl2002/acl02.tex
- Added pcfgparser to parser modules - Wrote advanced projects section - Added acknowledgements section (still needs to be filled in!)
svn/trunk@616

% To do:
%   - Write up other approaches
%   - Write up advanced projects section

@@

% [EL] ``punitive'' like a strange word.
% [EL] I couldn't find a REF for ``executable pseudocode'' -- closest
% I got was http://www.thinkware.se/cgi-bin/thinki.cgi/PythonQuotes
% [SB] changed Python citation to URL
In surveying the available languages, we believe that Python offers an
especially good fit to the above requirements.  Python is an
object-oriented scripting language developed by Guido van Rossum
and available on all platforms (\url{www.python.org}).  Python offers a
shallow learning curve; it was designed to be easily learnt by
children \cite{rossum99}.  As an interpreted language, Python is
suitable for rapid prototyping.  Python code is exceptionally
readable, and it has been praised as ``executable pseudocode.''
Python is an object-oriented language, but not punitively so, and it
is easy to encapsulate data and methods inside Python classes.
Finally, Python has an interface to the Tk graphics toolkit
\cite{tkinter}, and writing graphical interfaces is straightforward.

@@

% [EL] I commented out the list of documentation types, since we have
% a section that talks about the documentation.
\paragraph{\textit{Documentation.}} The toolkit, its data structures,
and its implementation all need to be carefully and thoroughly
documented.  
%Three documentation types are necessary: tutorials, to
%teach students how to use the toolkit to perform specific tasks;
%technical reports, to explain the toolkit's design and implementation;
%and reference documentation, to describe every module, interface,
%class and method.  
All nomenclature must be carefully chosen and consistently used.

@@

Three modules provide implementations for these abstract interfaces.
The \texttt{srparser} module implements a simple shift-reduce parser.
The \texttt{chartparser} module provides a flexible parser that uses a
\emph{chart} to record hypotheses about syntactic constituents.  And
the \texttt{rechunkparser} module defines a transformational
regular-expression based implementation of the chunk parser interface.
% [EL] Add the PCFG parser(s) to this once I'm done with my 630 project

@@

% [EL] This still needs work...
The visualization modules provide interfaces for interaction and
experimentation; they do not directly implement any NLP data
structures or tasks.  Simplicity of implementation is therefore less
of an issue for the visualization modules than it is for the rest of
the toolkit.

@@

% [EL] Provide some indication of how much tutorial material is
% available?  (Currently, 8 tutorials, mean=16pp, range=6pp-52pp.)
\paragraph{\textit{Tutorials}} teach students how to use the toolkit,
in the context of performing specific tasks.  Each tutorial focuses on
a single domain, such as tagging, probabilistic systems, or text
classification.  The tutorials include a high-level discussion that
explains and motivates the domain, followed by a detailed
walk-through that uses examples to show how NLTK can be used to
perform specific tasks.

@@

Advanced projects typically involve the development of entirely new
functionality for a previously unsupported NLP task, or the
development of a complete system out of existing and new modules.

% [EL] FILL THIS IN!!!!

@@

% [EL] Add something saying explicitly that there's still a lot to
% do?  (In response to reviewer comment)
%Future work: 
We plan to continue extending the breadth of materials covered by the
toolkit.  We are currently working on NLTK modules for Hidden Markov
Models, probabilistic parsing, and tree adjoining grammars.  We also
plan to increase the number of algorithms implemented by some existing
modules, such as the text classification module.

@@

@@

\end{document}

>>>>

% To do:
%   - Add acknowledgements

@@

% [EL] ``punitive'' seems like a strange word.
% [EL] I couldn't find a REF for ``executable pseudocode'' -- closest
% I got was http://www.thinkware.se/cgi-bin/thinki.cgi/PythonQuotes
% [SB] changed Python citation to URL
In surveying the available languages, we believe that Python offers an
especially good fit to the above requirements.  Python is an
object-oriented scripting language developed by Guido van Rossum
and available on all platforms (\url{www.python.org}).  Python offers a
shallow learning curve; it was designed to be easily learnt by
children \cite{rossum99}.  As an interpreted language, Python is
suitable for rapid prototyping.  Python code is exceptionally
readable, and it has been praised as ``executable pseudocode.''
Python is an object-oriented language, but not punitively so, and it
is easy to encapsulate data and methods inside Python classes.
Finally, Python has an interface to the Tk graphics toolkit
\cite{tkinter}, and writing graphical interfaces is straightforward.

@@

% [EL] I commented out the list of documentation types, since we have
% a section that talks about the documentation.
\paragraph{\textit{Documentation.}} The toolkit, its data structures,
and its implementation all need to be carefully and thoroughly
documented.  All nomenclature must be carefully chosen and
consistently used.

@@

Four modules provide implementations for these abstract interfaces.
The \texttt{srparser} module implements a simple shift-reduce parser.
The \texttt{chartparser} module defines a flexible parser that uses a
\emph{chart} to record hypotheses about syntactic constituents.  The
\texttt{pcfgparser} module provides a variety of different parsers for
probabilistic grammars.  And the \texttt{rechunkparser} module defines
a transformational regular-expression based implementation of the
chunk parser interface.

@@

% [EL] This still needs work...
The visualization modules provide interfaces for interaction and
experimentation; they do not directly implement NLP data structures or
tasks.  Simplicity of implementation is therefore less of an issue for
the visualization modules than it is for the rest of the toolkit.

@@

% [EL] Provide some indication of how much tutorial material is
% available?  (Currently, 9 tutorials, mean=16pp, range=6pp-52pp.)
\paragraph{\textit{Tutorials}} teach students how to use the toolkit,
in the context of performing specific tasks.  Each tutorial focuses on
a single domain, such as tagging, probabilistic systems, or text
classification.  The tutorials include a high-level discussion that
explains and motivates the domain, followed by a detailed
walk-through that uses examples to show how NLTK can be used to
perform specific tasks.

@@

NLTK provides students with a flexible framework for advanced
projects.  Typical projects involve the development of entirely new
functionality for a previously unsupported NLP task, or the
development of a complete system out of existing and new modules.

% Broad coverage 
The toolkit's broad coverage allows students to explore a wide variety
of topics.  In our introductory computational linguistics course,
topics for student projects included text generation, word sense
disambiguation, collocation analysis, and morphological analysis.

% Less grunt-work
NLTK eliminates the tedious infrastructure-building that is typically
associated with advanced student projects, by providing students with
the basic data structures, tools, and interfaces that they need.  This
allows the students to concentrate on the problems that interest them.

% Meaningful contribution
The collaborative, open-source nature of the toolkit can provide
students with a sense that their projects are meaningful
contributions, and not just exercises.  Several of the students in our
course have expressed interest in incorporating their projects into
the toolkit.

% Good examples
Finally, many of the modules included in the toolkit provide students
with good examples of what projects should look like, with well
thought out interfaces, clean code structure, and thorough
documentation.

\subsubsection*{Example: Probabilistic Parsing}
% [EL] I'm a bit uncomfortable including this description without
% mentioning that I'm the author -- seems a little misleading.  But at
% the same time, it seems like it's much less interesting if I say
% that I wrote it, given that I wrote the majority of the toolkit
% anyway.  :-/  Perhaps we should just leave out this section.  Any
% ideas?

The probabilistic parsing module was created as a class project for a
statistical NLP course.  The toolkit provided the basic data types and
interfaces for parsing.  The project extended these, adding a new
probabilistic parsing interface, and using subclasses to create a
probabilistic version of the CFG data structure.  These new components
were used in conjunction with several existing components, such as the
chart data structure, to define two implementations of the
probabilistic parsing interface.  Finally, a tutorial was written that
explained the basic motivations and concepts behind probabilistic
parsing, and described the new interfaces, data structures, and
parsers.

@@

% [EL] Add something saying explicitly that there's still a lot to
% do?  (In response to reviewer comment)
%Future work: 
We plan to continue extending the breadth of materials covered by
the toolkit.  We are currently working on NLTK modules for Hidden
Markov Models, language modeling, and tree adjoining grammars.  We
also plan to increase the number of algorithms implemented by some
existing modules, such as the text classification module.

@@

% ===================== Acknowledgements =====================
\section{Acknowledgments}
\label{sec:acknowledgments}

\emph{Add acknowledgements: reviewers; funders; students of 530;
others?}

@@

\end{document}
% LocalWords:  stevenbird Exp pcfgparser subclasses CFG DCG HPSG toolkits CMU
